\singlespacing \vspace{-2.5cm} \LARGE \textbf{Trade-offs among plant reproductive traits determine interactions with floral visitors}

### 2021-10-05 Manuscript/Manuscript.Rmd
edits

__Plant life strategies are often delimited by vegetative and physiological traits but little is known about how floral and reproductive biology traits drive these strategies, and in turn shape plant interactions with floral visitors. Here, we compiled 13 floral, 4 reproductive biology and 3 vegetative traits for 1,506 plant species from 28 plant-pollinator network studies across 18 different countries. We investigated the associations among these traits, pollinator visitation and the functional role of plant species within the networks (interaction frequency, normalized degree and specialization). We found that 51.8% of trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size (PC1, 26.72%). The second axis indicated a trade-off for the level of pollinator dependency (PC2, 25.08%). Although the main axes of trait variation did not fully explain pollinator visitation rates, different plant life strategies were associated with visitation rates and pollinator functional groups. Overall, the main traits that determined plant species’ functional roles were height, nectar concentration, pollen grains per flower, number of ovules, style length, selfing level and flower width. Our results highlight the need to consider plant reproductive and floral traits to improve understanding of plant life strategies and plant-pollinator interactions at broader spatial scales.__

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10,fig.pos = "t",fig.cap= "**Fig. 1 | Plant life-history strategies.** Phylogenetically informed principal component analysis (pPCA) of 1,236 plant species from 28 plant-pollinator network studies. The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) across the two main axes of trait variation. The length of the arrows indicate the weight of the variables on each principal component and the dashed lines show the opposed direction of trait variation. The icons at both ends of arrows and dashed lines illustrate the extreme form of the trait continuum."}

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=12, fig.width=18,fig.pos = "!H",fig.cap= "**Fig. 2 | Location of the different qualitative traits on the trait space.** The panel is composed by the traits that showed statistical association with the first two axes of trait variation: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f)."}

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.pos = "!H",fig.cap= "**Fig. 3 | Interaction (yes/no) and visitation rates across the three main axes of trait variation per floral visitor guild.** Fitted posterior estimates of the presence/absence of interaction (a, b and c) and number of visits (d, e and f) made by the different floral visitors guilds in relation to PC1, PC2 and PC3. PC1 represents the flower number - flower size trade-off, PC2 represents the pollinator dependence trade-off and PC3, the style length trade-off. For visualization purposes, due to large differences between the visitation rates of bees and the rest of guilds, the number of visits was log-transformed (Y-axis of lower panel)."}

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=4.25, fig.width=12,fig.pos = "!H",fig.cap= "**Fig. 4 | Contribution of traits in plant's network roles.** Regression tree analysis of interaction frequency (log-transformed), normalized degree and specialization for the subset of species with quantitative data for pollen and nectar traits. The superior value inside the node indicates the mean value of the different species-level metric and the lower value, the percentage of species that are considered in each node. Thus, the top node has the mean value of the named trait for the 100% of species. Each node has a yes/no question and when the condition is fulfilled, the branch turns to the ‘yes’ direction and when not, to the ‘no’ direction. This rationale is followed in all the regression trees as indicated in the first branch division of the topmost node of each tree."}

>>>>

__Plant life strategies are constrained by cost-benefit trade-offs that determine plant form and function. Despite the recent advances in the understanding of plant trade-offs with vegetative and physiological traits, little is known about the plant reproductive economics and how they constrain plant life strategies and shape interactions with floral visitors. Here, we investigated plant reproductive trade-offs and their association with floral visitors with a dataset of 1506 plant species with 17 reproductive and 3 vegetative traits from 28 plant-pollinator network studies across 18 countries. Then, we tested whether a species reproductive strategy predicted plant-pollinator interactions (presence-absence and visitation rate) and if the interaction strategy (its role in a pollination network) was predicted by individual traits. We found that 51.8% of trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size (PC1, 26.72%) and the second axis indicated a trade-off for the level of pollinator dependency (PC2, 25.08%). The reproductive trade-offs explained some of the qualitative interaction with pollinators (presence-absence) but none of the variability of visitation rates. However, noteworthy differences across pollinator guilds were found in both cases. Finally, we found that specific traits with special relevance of floral rewards were key in the understanding of the species network role. Our results highlight the need to consider plant reproductive trade-offs to improve our understanding of plant life strategies and plant-pollinator interactions at broader spatial scales.__

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10,fig.align='center',fig.width=10,fig.cap= "**Fig. 1 | Plant life-history strategies.** Phylogenetically informed principal component analysis (pPCA) of 1,236 plant species from 28 plant-pollinator network studies. The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) across the two main axes of trait variation. The length of the arrows indicate the weight of the variables on each principal component and the dashed lines show the opposed direction of trait variation. The icons at both ends of arrows and dashed lines illustrate the extreme form of the trait continuum."}

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=12, fig.width=18,fig.cap= "**Fig. 2 | Location of the different qualitative traits on the trait space.** The panel is composed by the traits that showed statistical association with the first two axes of trait variation: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f)."}

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.cap= "**Fig. 3 | Interaction (yes/no) and visitation rates across the three main axes of trait variation per floral visitor guild.** Fitted posterior estimates of the presence/absence of interaction (a, b and c) and number of visits (d, e and f) made by the different floral visitors guilds in relation to PC1, PC2 and PC3. PC1 represents the flower number - flower size trade-off, PC2 represents the pollinator dependence trade-off and PC3, the style length trade-off. For visualization purposes, due to large differences between the visitation rates of bees and the rest of guilds, the number of visits was log-transformed (Y-axis of lower panel)."}

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=4.25, fig.width=12,fig.cap= "**Fig. 4 | Contribution of traits in plant's network roles.** Regression tree analysis of interaction frequency (log-transformed), normalized degree and specialization for the subset of species with quantitative data for pollen and nectar traits. The superior value inside the node indicates the mean value of the different species-level metric and the lower value, the percentage of species that are considered in each node. Thus, the top node has the mean value of the named trait for the 100% of species. Each node has a yes/no question and when the condition is fulfilled, the branch turns to the ‘yes’ direction and when not, to the ‘no’ direction. This rationale is followed in all the regression trees as indicated in the first branch division of the topmost node of each tree."}

### 2021-10-05 Manuscript/Manuscript.tex
edits

\textbf{Plant life strategies are often delimited by vegetative and
physiological traits but little is known about how floral and
reproductive biology traits drive these strategies, and in turn shape
plant interactions with floral visitors. Here, we compiled 13 floral, 4
reproductive biology and 3 vegetative traits for 1,506 plant species
from 28 plant-pollinator network studies across 18 different countries.
We investigated the associations among these traits, pollinator
visitation and the functional role of plant species within the networks
(interaction frequency, normalized degree and specialization). We found
that 51.8\% of trait variation was explained by two independent axes
that encompassed plant form and function. Specifically, the first axis
indicated the presence of a trade-off between flower number and flower
size (PC1, 26.72\%). The second axis indicated a trade-off for the level
of pollinator dependency (PC2, 25.08\%). Although the main axes of trait
variation did not fully explain pollinator visitation rates, different
plant life strategies were associated with visitation rates and
pollinator functional groups. Overall, the main traits that determined
plant species' functional roles were height, nectar concentration,
pollen grains per flower, number of ovules, style length, selfing level
and flower width. Our results highlight the need to consider plant
reproductive and floral traits to improve understanding of plant life
strategies and plant-pollinator interactions at broader spatial scales.}

@@

Here, we explore the potential trade-offs among plant floral and
reproductive biology traits and how these influence the structure of
plant-pollinator networks. First, we identify the major axes of trait
variation and trade-offs that determine plant form and function. Second,
we investigate how plant species' position in trait-space influences
their interaction strength with different guilds of floral visitors.
Finally, we investigate how the main axes of trait variation and
individual traits influence plant species roles within networks.

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}\textbf{Fig. 1 \textbar{} Plant
life-history strategies.} Phylogenetically informed principal component
analysis (pPCA) of 1,236 plant species from 28 plant-pollinator network
studies. The solid arrows indicate the direction of the different
quantitative traits (flower number, plant height, style length, flower
size, ovule number and level of autonomous selfing) across the two main
axes of trait variation. The length of the arrows indicate the weight of
the variables on each principal component and the dashed lines show the
opposed direction of trait variation. The icons at both ends of arrows
and dashed lines illustrate the extreme form of the trait continuum.}
\end{figure}

@@

\textbf{Network analyses.} Analyses were conducted on the subset of 60
weighted networks with interaction frequency sampled in a unique
flowering season and site, which included 556 plant and 1,126 pollinator
species. In total, our weighted network dataset (excluding meta-webs)
included 2,256 interactions of Anthophila-Hymenoptera (i.e., bees) with
plants, 1,768 non-Syrphidae-Diptera interactions, 845 Syrphidae
interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions
and 362 non-Anthophila-Hymenoptera interactions. Although floral
visitors are not always pollinators and the frequency of visits does not
consider each pollinator species
efficiency\textsuperscript{\protect\hyperlink{ref-ballantyne2015}{57}},
visitation rate provides valuable information of the contribution of
floral visitors to
pollination\textsuperscript{\protect\hyperlink{ref-vazquez2005}{58},\protect\hyperlink{ref-vazquez2012}{59}}.
Sampling methods varied across networks but this was accounted for in
analyses by considering them in the random effects of the modelling
process. All analyses were conducted in R version \emph{4.0.3}.

\textbf{Visitation patterns.} We used Bayesian modelling (see below for
details) to explore the effect of floral visitor groups and the main
axes of trait variation (pPCA with imputed dataset) on both qualitative
(presence/absence) and quantitative (visitation rate) floral
interactions per plant species. For this, we divided floral visitors
into six main guilds that differ in life form, behaviour and are likely
to play a similar ecological role: (i) Hymenoptera-Anthophila (bees),
(ii) Hymenoptera-non-Anthophila (non-bee Hymenoptera), (iii)
Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera and (vi)
Coleoptera. Moreover, because Hymenoptera-Anthophila was the most
represented group with 2,256 records and had the highest frequency of
visits of all groups, we also explored the presence-absence of
interaction and visitation rate of the main Hymenoptera-Anthophila
families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae)
on the trait space. In addition, we found that \emph{Apis mellifera} was
the floral visitor with the largest proportion of records counted
(7.55\% of the total). This finding is consistent with previous research
showing that \emph{A. mellifera} was the most frequent floral visitor in
a similar dataset of 80 plant-pollinator networks in natural
ecosystems\textsuperscript{\protect\hyperlink{ref-hung2018}{60}}. Hence,
to control for the effect of \emph{A. mellifera} on the observed
visitation patterns of bees, we conducted an analogous analysis
excluding \emph{A. mellifera}. We found that \emph{A. mellifera}, was
partly driving some of the observed trends on PC1 (Supplementary Fig.
S3). However, we did not detect major differences on PC2 and PC3.

>>>>

\textbf{Plant life strategies are constrained by cost-benefit trade-offs
that determine plant form and function. Despite the recent advances in
the understanding of plant trade-offs with vegetative and physiological
traits, little is known about the plant reproductive economics and how
they constrain plant life strategies and shape interactions with floral
visitors. Here, we investigated plant reproductive trade-offs and their
association with floral visitors with a dataset of 1506 plant species
with 17 reproductive and 3 vegetative traits from 28 plant-pollinator
network studies across 18 countries. Then, we tested whether a species
reproductive strategy predicted plant-pollinator interactions
(presence-absence and visitation rate) and if the interaction strategy
(its role in a pollination network) was predicted by individual traits.
We found that 51.8\% of trait variation was explained by two independent
axes that encompassed plant form and function. Specifically, the first
axis indicated the presence of a trade-off between flower number and
flower size (PC1, 26.72\%) and the second axis indicated a trade-off for
the level of pollinator dependency (PC2, 25.08\%). The reproductive
trade-offs explained some of the qualitative interaction with
pollinators (presence-absence) but none of the variability of visitation
rates. However, noteworthy differences across pollinator guilds were
found in both cases. Finally, we found that specific traits with special
relevance of floral rewards were key in the understanding of the species
network role. Our results highlight the need to consider plant
reproductive trade-offs to improve our understanding of plant life
strategies and plant-pollinator interactions at broader spatial scales.}

@@

Here, we explore the potential trade-offs among plant floral and
reproductive biology traits and how these influence the structure of
plant-pollinator networks. First, we identify the major axes of trait
variation and trade-offs that determine plant form and function. Second,
we investigate how plant species' position in trait-space influences
their interaction patterns with different guilds of floral visitors.
Finally, we investigate how the main axes of trait variation and
individual traits influence plant species roles within networks.

@@

\begin{figure}[H]

{\centering \includegraphics{output/figures/unnamed-chunk-1-1}

}

\caption{**Fig. 1 | Plant life-history strategies.** Phylogenetically informed principal component analysis (pPCA) of 1,236 plant species from 28 plant-pollinator network studies. The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) across the two main axes of trait variation. The length of the arrows indicate the weight of the variables on each principal component and the dashed lines show the opposed direction of trait variation. The icons at both ends of arrows and dashed lines illustrate the extreme form of the trait continuum.}\label{fig:unnamed-chunk-1}
\end{figure}

@@

\textbf{Network analyses.} Analyses were conducted on the subset of 60
weighted networks sampled in a unique flowering season and site, which
included 556 plant and 1,126 pollinator species. These networks were
analysed in their qualitative (presence-absence) and quantitative
(interaction frequency) form. First, we analysed the binary version of
these weighted networks with presence-absence information that assumes
equal weight across interactions. Second, we analysed the untransformed
weighted networks with interaction frequency that accounts for the
intensity of the interaction. Although floral visitors are not always
pollinators and interaction frequency does not consider each pollinator
species
efficiency\textsuperscript{\protect\hyperlink{ref-ballantyne2015}{57}},
interaction frequency can provide valuable information of the
contribution of floral visitors to
pollination\textsuperscript{\protect\hyperlink{ref-vazquez2005}{58},\protect\hyperlink{ref-vazquez2012}{59}}.
In total, our network dataset (excluding meta-webs and non-weighted
networks) included 2,256 interactions of Anthophila-Hymenoptera (i.e.,
bees) with plants, 1,768 non-Syrphidae-Diptera interactions, 845
Syrphidae interactions, 437 Lepidoptera interactions, 432 Coleoptera
interactions and 362 non-Anthophila-Hymenoptera interactions. Sampling
methods varied across networks but this was accounted for in analyses by
considering them in the random effects of the modelling process. All
analyses were conducted in R version \emph{4.0.3}.

\textbf{Visitation patterns.} We used Bayesian modelling (see below for
details) to explore the effect of floral visitor groups and the main
axes of trait variation (pPCA with imputed dataset) on both qualitative
(presence/absence) and quantitative (visitation rate) floral
interactions per plant species. For this, we divided floral visitors
into six main guilds that differ in life form, behaviour and are likely
to play a similar ecological role: (i) Hymenoptera-Anthophila (bees),
(ii) Hymenoptera-non-Anthophila (non-bee Hymenoptera), (iii)
Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera and (vi)
Coleoptera. Moreover, because Hymenoptera-Anthophila was the most
represented group with 2,256 records and had the highest frequency of
visits of all groups, we also explored the presence-absence of
interaction and visitation rate of the main Hymenoptera-Anthophila
families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae)
on the trait space. In addition, we found that \emph{Apis mellifera} was
the floral visitor with the largest proportion of records counted
(7.55\% of the total). This finding is consistent with previous research
showing that \emph{A. mellifera} was the most frequent floral visitor in
a similar dataset of 80 plant-pollinator networks in natural
ecosystems\textsuperscript{\protect\hyperlink{ref-hung2018}{60}}. Hence,
to control for the effect of \emph{A. mellifera} on the observed
visitation patterns of bees, we conducted an analogous analysis with
presence-absence of interaction and visitation rate excluding \emph{A.
mellifera}. We found that \emph{A. mellifera}, was partly driving some
of the observed trends on PC1 (Supplementary Fig. S3). However, we did
not detect major differences on PC2 and PC3.

### 2021-10-05 Supplementary_Material/Supplementary_Material.Rmd
edits

---
title: \singlespacing \vspace{-2.5cm} \LARGE \textbf{Trade-offs among plant floral and reproductive traits determine interactions with floral visitors}

@@

```{r pressure, echo=FALSE, fig.cap="", out.width = '65%',fig.align='center', fig.cap="\\textbf{Supplementary text Fig. 1 | Main floral morphologies considered in this study.} All species were groupped within these categories for analysis: tube (a), papilonaceous (b), open (c), capitulum (d), campanulate (e) and brush (f)."}
knitr::include_graphics("../Images/flower_shapes.png")
```

>>>>

---
title: \singlespacing \vspace{-2.5cm} \LARGE \textbf{Trade-offs among plant reproductive traits determine interactions with floral visitors}

@@

```{r pressure, echo=FALSE, fig.cap="", out.width = '65%',fig.align='center', fig.cap="\\textbf{Supplementary text Fig. 1 | Main floral morphologies considered in this study.} All species were groupped within these categories for analysis: (a) tube, (b) papilonaceous, (c) open, (d) capitulum, (e) campanulate and (f) brush."}
knitr::include_graphics("../Images/flower_shapes.png")
```
